---
title: "DCS375_Final_Project"
author: "Harrison Hartnett"
date: "`r Sys.Date()`"
output: html_document
---

## Abstract:

The purpose of this script is to create a network visualization of the private security database (PSD) with emphasis on which countries tend to employ the most private military/security companies, and where they tend to deploy them.

## Checking for and Installing Packages
We are going to start by identifying any packages needed for our analysis. In order to avoid any unnecessary re-installation or loading, we are going to set up a way to automatically check and then load only what we need.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Store string containing all required packages
my_packages <- c('igraph', 'RColorBrewer', 'ggplot2', 'dplyr', 'janitor', 'stringr', 'stopwords', 'bipartite')

```

Figure out which of these packages is already installed
``` {r}
# Store all installed packages
ya_installed <- library()$results[,1]

# Check whether required packages are already installed and grab only those that still need installation
need_install<-my_packages[!(my_packages %in% ya_installed)]

#install required packages
lapply(need_install, install.packages, character.only = TRUE)

```

Now, load only unloaded packages
``` {r}
# Store all installed packages
ya_loaded <- (.packages())

# Check whether required packages are already installed and grab only those that still need installation
need_load<-my_packages[!(my_packages %in% ya_loaded)]

# Load required packages
lapply(need_load, require, character.only = TRUE)
```
## Import the Private Security Database
```{r}
PSD <- read.csv("PSD_V4_1990-2007_csv.csv", row.names=NULL, sep=";")

View(PSD)
head(PSD)
summarize(PSD)
```

# Preprocess the data
(To be converted into a matrix and graph object)
```{r}
prepped_data <- PSD %>%
  group_by(clientsc, locsc) %>%#I want to know which countries are sending PMSCs where
  summarize(
    contract_no = sum(nofirm)#Summarize the number of PMSCs deployed to a specific country
  )
View(prepped_data)

#Change country codes to actual names
for (i in prepped_data) {
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "ANG", "Angola")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "AUL", "Australia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "AZE", "Azerbaijan")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "BOS", "Bosnia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "CAN", "Canada")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "COL", "Columbia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "CON", "Congo")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "CPA", "Commonwealth Parlimentary Association")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "CRO", "Croatia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "DRC", "Democratic Republic of Congo")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "ETH", "Ethiopia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "IRQ", "Iraq")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "ITA", "Italy")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "LBR", "Liberia")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "MZM", "Mozambique")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "PHI", "Philipines")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "RWA", "Rwanda")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "SAF", "South Africa")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "SIE", "Sierra Leone")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "UGA", "Uganda")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "UK", "United Kingdom")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "UKG", "United Kingdom")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "UN", "United Nations")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "USA", "United States")
  prepped_data$clientsc <- str_replace_all(prepped_data$clientsc, "WB", "West Bank")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "ANG", "Angola")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "IRQ", "Iraq")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "COL", "Columbia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "AZE", "Azerbaijan")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "BOS", "Bosnia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "AFG", "Afghanistan")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "CON", "Congo")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "CRO", "Croatia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "DRC", "Democratic Republic of Congo")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "ETH", "Ethiopia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "LBR", "Liberia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "MZM", "Mozambique")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "PHI", "Philipines")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "RWA", "Rwanda")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "BUI", "Burundi")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "SIE", "Sierra Leone")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "UGA", "Uganda")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "YUG", "Yugoslavia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "SUD", "Sudan")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "NEP", "Nepal")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "SOM", "Somalia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "IVO", "Ivory Coast")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "GRG", "Georgia")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "NIG", "Nigeria")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "ALB", "Albania")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "GUA", "Guatemala")
  prepped_data$locsc <- str_replace_all(prepped_data$locsc, "PER", "Peru")
}
View(prepped_data)
```

# Create a graph object that we can visualize
```{r}
PSD_matrix <- as.matrix(prepped_data)
#PSD_el <- as.edgelist(prepped_data)
View(PSD_matrix)

PSD.g <- graph_from_data_frame(PSD_matrix, directed = TRUE) #Directed because I want to know which country a contract is coming from
PSD.g #and where it is going
```

# Assign Node and Edge attributes
```{r}
V(PSD.g)$vertex_degree <- igraph::degree(PSD.g) #Assign a degree attribute

edge_weights <- E(PSD.g)$edge_weight <- prepped_data$contract_no #Make edgeweights the number of firms employed

#Preliminary Visualization
set.seed(375)#For reproduceability
lo_nice = layout_nicely(PSD.g)#layout_nicely automatically selects the best layout
png(file="PrivateSecurityNetwork.png", width=12000, height=10000, res = 400)#Save the visualization as a png
plot(PSD.g, layout = lo_nice, vertex.size = log(V(PSD.g)$vertex_degree) * 5, edge.width = (log(edge_weights) + 1) * 4, edge.arrow.size = 0.9,
     vertex.label.color = "black", vertex.label.cex = 3) +
  title("Private Military Contracts by Country", cex.main = 4)
  #+ palette.colors(palette = "RdPu")#cex.main = size of title
dev.off()#Necessary when saving the visualization as a png
```

# Calculate the 5 number summary
```{r}
print("PSD.g 5 number summary")
print("Size:")
vcount(PSD.g)
print("Density")
edge_density(PSD.g)
print("Components:")
components(PSD.g)
print("Diameter:")
diameter(PSD.g)
print("Transitivity:")
transitivity(PSD.g)
```

# Creating a Bipartite Graph
```{r}
#Build an edgelist
prepped_data2 <- PSD %>%
  group_by(clientsc) %>%
  summarize(
    locsc = locsc
  )
for (i in prepped_data) {
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "ANG", "Angola")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "AUL", "Australia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "AZE", "Azerbaijan")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "BOS", "Bosnia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "CAN", "Canada")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "COL", "Columbia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "CON", "Congo")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "CPA", "Commonwealth Parlimentary Association")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "CRO", "Croatia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "DRC", "Democratic Republic of Congo")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "ETH", "Ethiopia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "IRQ", "Iraq")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "ITA", "Italy")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "LBR", "Liberia")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "MZM", "Mozambique")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "PHI", "Philipines")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "RWA", "Rwanda")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "SAF", "South Africa")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "SIE", "Sierra Leone")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "UGA", "Uganda")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "UK", "United Kingdom")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "UKG", "United Kingdom")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "UN", "United Nations")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "USA", "United States")
  prepped_data2$clientsc <- str_replace_all(prepped_data2$clientsc, "WB", "West Bank")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "ANG", "Angola")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "IRQ", "Iraq")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "COL", "Columbia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "AZE", "Azerbaijan")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "BOS", "Bosnia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "AFG", "Afghanistan")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "CON", "Congo")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "CRO", "Croatia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "DRC", "Democratic Republic of Congo")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "ETH", "Ethiopia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "LBR", "Liberia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "MZM", "Mozambique")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "PHI", "Philipines")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "RWA", "Rwanda")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "BUI", "Burundi")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "SIE", "Sierra Leone")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "UGA", "Uganda")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "YUG", "Yugoslavia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "SUD", "Sudan")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "NEP", "Nepal")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "SOM", "Somalia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "IVO", "Ivory Coast")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "GRG", "Georgia")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "NIG", "Nigeria")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "ALB", "Albania")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "GUA", "Guatemala")
  prepped_data2$locsc <- str_replace_all(prepped_data2$locsc, "PER", "Peru")
}
#View(prepped_data2)

#prepped_data2.g <- graph_from_biadjacency_matrix(prepped_data2, weighted = TRUE)
#edge.prepped_data2 <- as_edgelist(prepped_data2.g)

webID <- data.frame(matrix("privatecontract", nrow = nrow(prepped_data2), ncol = 1))
new_obj <- cbind(prepped_data2, webID)
colnames(new_obj)<- c("clientsc", "locsc", "WebID")
View(new_obj)
```

# Creating a bipartite network object
To analyze the degree distribution and node specific summary statistics
```{r}
new_obj_df <- as.data.frame(new_obj)
private_vis1 <- frame2webs(new_obj_df, varnames = c("clientsc", "locsc", "WebID"),
                           type.out = "list", emptylist = TRUE)#Output type = list

#Finding the degree distribution
degreedistr(private_vis1$'privatecontract', plot.it=TRUE, pure.call=TRUE, silent=TRUE, level="both")
```

# Summaries
```{r}
#Network Level Summary
networklevel <- networklevel(private_vis1$'privatecontract')

#Node Level Summaries
node_summary <- specieslevel(private_vis1$'privatecontract') #Specieslevel can calculate summary statistics for individual nodes

#Specific stasitics: I chose the Unites States
node_summary$'higher level'['United States',] #Summarize stats for the U.S node
node_summary$'lower level'['United Nations',] #Summarize stats for the U.N node

networklevel
node_summary
```
It makes sense that there isn't a great fit for the givers
The truncated power law means you have few countries are contributing a lot - and you have many countries who do very little
Linked for interpretation of these results